Chord: A Versatile Platform for Program Analysis

### 2011-04-04 main/doc/getting_started.tex
documentation

\chapter{Getting Started}
\label{chap:getting-started}

@@

To obtain Chord's pre-built binaries, download file \xlink{chord-bin-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-bin-2.0.tar.gz}.
It contains the following files:

@@

To obtain Chord's source code, download the following files:

\begin{itemize}
\item
Mandatory: file \xlink{chord-src-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-src-2.0.tar.gz}, which contains Chord's source code and the jars of libraries used by Chord.
\item
Optional: file \xlink{chord-libsrc-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-libsrc-2.0.tar.gz}, which contains the source
code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.)
\end{itemize}

Alternatively, you can obtain the latest development snapshot from the SVN repository:

\begin{verbatim}
    prompt> svn checkout http://jchord.googlecode.com/svn/trunk/ chord
\end{verbatim}

@@

\begin{itemize}
\item
\code{main/} contains Chord's source code and the jars of libraries used by Chord.
\item
\code{libsrc/} contains the source code of libraries used by Chord (e.g., joeq, javassist, bddbddb, etc.).
\item
\code{test/} contains Chord's regression tests.
\item
many more; these might eventually move into \code{main/}.
\end{itemize}

@@

\begin{itemize}
\item
A JVM with JDK 5 or higher, e.g. \xlink{IBM J9}{http://www.ibm.com/developerworks/java/jdk/} or
\xlink{Oracle HotSpot}{http://www.oracle.com/technetwork/java/javase/downloads/index.html}.
\item
\xlink{Apache Ant}{http://ant.apache.org/}, a Java build tool.
\end{itemize}

Directory {\tt main/} contains a file named {\tt build.xml} which is
interpreted by Apache Ant.  To see the various targets available, run
the following command in that directory:

\begin{verbatim}
    prompt> ant
\end{verbatim}

To compile Chord, run the following command in the same directory:

\begin{verbatim}
    prompt> ant compile
\end{verbatim}

This will compile Chord's Java sources from \code{main/src/} to class files
in \code{main/classes/}, as well as build a jar file \code{main/chord.jar} containing
these class files as well as the class files in the jars of libraries that are used by Chord
and are provided under \code{main/lib/} (e.g., \code{joeq.jar}, \code{javassist.jar},
\code{bddbddb.jar}, etc.).
Additionally:

@@

If system property \code{chord.use.buddy} is set to \code{true},
then the C source code of BDD library \xlink{BuDDy}{http://buddy.sourceforge.net/} 
from \code{main/bdd/} will be compiled to a shared library in \code{main/}
(\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and
\code{libbuddy.dylib} on MacOS); this library is used 
by BDD-based Datalog solver \xlink{bddbddb}{http://bddbddb.sourceforge.net/} 
in Chord for running analyses written in Datalog.

@@

If system property \code{chord.use.jvmti} is set to \code{true},
then the C++ source code of the JVMTI-based bytecode instrumentation agent from
\code{main/agent/} will be compiled to a shared library in \code{main/}
(\code{libchord_instr_agent.so} on all architectures); this agent is
used in Chord for computing analysis scope dynamically and for running
dynamic analyses.
\end{itemize}

Properties \code{chord.use.buddy} and \code{chord.use.jvmti} are defined in file \code{main/chord.properties}.  The
default value of both these properties is \code{false}.
If you set either of them to \code{true}, then you will also need a utility like GNU Make (to run the
\code{Makefile}'s under \code{main/bdd/} and \code{main/agent/}) and a C++ compiler.

@@

Running Chord requires a JVM with JDK 5 or higher.
There are two equivalent ways to run Chord.
One way is to run the following command in the \code{main/} directory:

\begin{verbatim}
    prompt> ant -D<key1>=<value1> ... -D<keyN>=<valueN> run
\end{verbatim}

The above requires Apache Ant (a Java build tool) to be installed on your machine.  The alternative, which
does not require Apache Ant, is to run the following command:

\begin{verbatim}
    prompt> java -cp <...>/chord.jar -D<key1>=<value1> ... -D<keyN>=<valueN> chord.project.Boot
\end{verbatim}

where \code{<...>} denotes the absolute or relative path of the directory containing file \code{chord.jar};
that directory is also expected to contain any other binaries in Chord's installation
(e.g., \code{libbuddy.so} and \code{libchord_instr_agent.so}).

Each ``\code{-D<key>=<value>}" argument above sets the system property named \code{<key>} to
the value denoted by \code{<value>}.  The only way to specify inputs to Chord is via system properties; there is no
command-line argument processing.
Chapter\ref{chap:properties} describes all system properties recognized by Chord.

>>>>

\xname{getting_started}
\chapter{Getting Started}
\label{chap:getting-started}

@@

To obtain Chord's pre-built binaries, download and uncompress file
\xlink{chord-bin-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-bin-2.0.tar.gz}.
It includes the following files:

@@

To obtain Chord's source code, download and uncompress the following
files:

\begin{itemize}
\item
Mandatory: file
\xlink{chord-src-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-src-2.0.tar.gz},
which contains Chord's source code and jars of libraries used by
Chord.
\item
Optional: file
\xlink{chord-libsrc-2.0.tar.gz}{http://jchord.googlecode.com/files/chord-libsrc-2.0.tar.gz},
which contains the source code of libraries used by Chord (e.g., joeq,
javassist, bddbddb, etc.)
\end{itemize}

Alternatively, you can obtain the latest development snapshot from the
SVN repository by running the following command:

\begin{framed}
\begin{verbatim}
svn checkout http://jchord.googlecode.com/svn/trunk/ chord
\end{verbatim}
\end{framed}

@@

\begin{itemize}
\item
\code{main/} contains Chord's source code and jars of libraries used by Chord.
\item
\code{libsrc/} contains the source code of libraries used by Chord
(e.g., joeq, javassist, bddbddb, etc.).
\item
\code{test/} contains Chord's regression tests.
\item
many more; these might eventually move under \code{main/}.
\end{itemize}

@@

\begin{itemize}
\item
JVM for Java 5 or higher, e.g.
\xlink{IBM J9}{http://www.ibm.com/developerworks/java/jdk/} or
\xlink{Oracle HotSpot}{http://www.oracle.com/technetwork/java/javase/downloads/index.html}.
\item
\xlink{Apache Ant}{http://ant.apache.org/}, a Java build tool.
\end{itemize}

Chord's main directory contains a file named {\tt build.xml} which is
interpreted by Apache Ant.  To see the various possible targets,
simply run command ``{\tt ant}" in that directory.

To compile Chord, run command ``{\tt ant compile}" in the same directory.
This will compile Chord's Java sources from \code{src/} to class
files in \code{classes/}, as well as build a jar file
\code{chord.jar} that contains these class files as well as those
in the jars of libraries that are used by Chord and are
provided under \code{lib/} (e.g., \code{joeq.jar},
\code{javassist.jar}, \code{bddbddb.jar}, etc.).  Additionally:

@@

If system property \code{chord.use.buddy} is set to \code{true}, then
the C source code of BDD library
\xlink{BuDDy}{http://buddy.sourceforge.net/} from directory \code{bdd/}
will be compiled to a shared library named
(\code{libbuddy.so} on Linux, \code{buddy.dll} on Windows, and
\code{libbuddy.dylib} on MacOS; this library is used by BDD-based
Datalog solver \xlink{bddbddb}{http://bddbddb.sourceforge.net/} in
Chord for running analyses written in Datalog.

@@

If system property \code{chord.use.jvmti} is set to \code{true}, then
the C++ source code of the JVMTI-based bytecode instrumentation agent
from directory \code{agent/} will be compiled to a shared library named
\code{libchord_instr_agent.so} on all architectures;
this agent is used in Chord for computing analysis scope dynamically
and for running dynamic analyses.
\end{itemize}

Properties \code{chord.use.buddy} and \code{chord.use.jvmti} are
defined in file \code{chord.properties} in Chord's main directory.
The default value of both these properties is \code{false}.  If you set
either of them to \code{true}, then you will also need a utility like
GNU Make (to run the \code{Makefile}'s in directories \code{bdd/} and
\code{agent/}) and a C++ compiler (to build the above shared libraries).

@@

Running Chord requires a JVM supporting Java 5 or higher.
There are two equivalent ways to run Chord.
One way, which is available only in the source installation of Chord, is to run the
following command in Chord's main directory:

\begin{framed}
\begin{verbatim}
ant -D<key1>=<val1> ... -D<keyN>=<valN> run
\end{verbatim}
\end{framed}

The above requires Apache Ant (a Java build tool) to be installed on
your machine.  The alternative, which does not require Apache Ant and is
available in both the source and binary installations of Chord, is
to run the following command:

\begin{framed}
\begin{verbatim}
java -cp <...>/chord.jar -D<key1>=<val1> ... -D<keyN>=<valN> chord.project.Boot
\end{verbatim}
\end{framed}

where \code{<...>} denotes the absolute or relative path of the
directory containing file \code{chord.jar}; that directory is also
expected to contain any shared libraries in Chord's installation (e.g.,
\code{libbuddy.so} and \code{libchord_instr_agent.so}).

Each ``\code{-D<key>=<val>}" argument above sets the system property
named \code{<key>} to the value denoted by \code{<val>}.  The only
way to specify inputs to Chord is via system properties; there is no
command-line argument processing.  Chapter \ref{chap:properties}
describes all system properties recognized by Chord.

### 2011-04-04 main/doc/program_representation.tex
documentation

Chord uses \xlink{Joeq}{http://joeq.sourceforge.net/} to translate
Java bytecode, one class file at a time, into a three-address-like
intermediate representation of the input Java program called {\it
quadcode}.  This chapter describes all aspects of quadcode and how it
relates to bytecode.  It first explains how to pretty-print bytecode
and quadcode (Section \ref{sec:pretty-printing}) which is useful for
debugging analyses and deciphering their output.  The remaining
sections describe the quadcode representation along with the API of
Joeq and Chord for navigating it.  Briefly, the representation
consists of a set of classes that may be loaded
(Section \ref{sec:program}).  Each class has a set of fields and
methods, collectively called members of the class
(Section \ref{sec:class-members}).  The representation of a concrete
method (Section \ref{sec:methods}) consists of a control-flow graph
(CFG).  The representation of a CFG (Section \ref{sec:cfgs}) consists
of a set of local variables and a set of basic blocks.  Each basic
block contains zero or more primitive statements called quads
(Section \ref{sec:quads}).  Finally, the most common way to traverse
all quads is discussed (Section \ref{sec:traversing}).

@@

The quadcode representation of each method is a unique object of
class \javadoc{joeq.Class.jq_Method}{joeq/Class/jq_Method.html}.  Components of the method (e.g.,
its control-flow graph) can be accessed using the following API provided
by that class.

@@

The control-flow graph (CFG) of each method consists of a set of local
variables of the method, called a ``register factory'', and a graph
whose nodes are basic blocks and whose edges denote flow of control
between basic blocks.

@@

The register factory contains one variable per argument of the method
and one variable per temporary declared in the method body.
Temporaries include those declared by programmers as well as those
generated by Joeq.  The reason for the latter is that the quadcode
representation is register-based and is constructed from Java
bytecode, which is stack-based; moreover, Joeq does the Static Single
Assignment (SSA) transformation by default, which introduces
temporaries to ensure that there is at most one static assignment to
any variable.

The register factory is a unique object
of class \javadoc{joeq.Compiler.Quad.RegisterFactory}{joeq/Compiler/Quad/RegisterFactory.html}.

@@

\begin{framed}
\begin{verbatim}
Operator
    |--- Move
    |--- Phi
    |--- Binary
    |--- Unary
    |--- New
    |--- NewAray
    |--- MultiNewArray
    |--- CheckCast
    |--- InstanceOf
    |--- ALength
    |--- ALoad
    |--- AStore
    |--- Getstatic
    |--- Putstatic
    |--- Getfield
    |--- Putfield
    |--- Return
    |--- Branch
    |       |--- IntIfCmp
    |       |--- Goto
    |       |--- Jsr
    |       |--- Ret
    |       |--- LookupSwitch
    |       |--- TableSwitch
    |--- Invoke
    |       |--- InvokeVirtual
    |       |--- InvokeStatic
    |       |--- InvokeInterface
    |--- Monitor
            |--- MONITORENTER
            |--- MONITOREXIT
\end{verbatim}
\end{framed}

@@

QuadVisitor qv = new QuadVisitor.EmptyVisitor() {
    public void visitMove(Quad q) { ... }
    public void visitPhi(Quad q) { ... }
    public void visitBinary(Quad q) { ... }
    ...
};
Program program = Program.g();
for (jq_Method m : program.getMethods()) {
    if (!m.isAbstract()) {
        ControlFlowGraph cfg = m.getCFG();
        ListIterator.BasicBlock it = cfg.reversePostOrderIterator();
        while (it.hasNext()) {
            BasicBlock b = it.nextBasicBlock();
            for (int i = 0; i < b.size(); i++) {
                Quad q = b.getQuad(i);
                q.accept(qv);
            }
        }
    }
}
\end{verbatim}
\end{framed}

>>>>

Chord uses \xlink{Joeq}{http://joeq.sourceforge.net/} to translate
Java bytecode, one class file at a time, into a three-address-like
intermediate representation of the input Java program called {\it
quadcode}.  This chapter describes all aspects of quadcode and how it
relates to bytecode.  It first explains how to pretty-print bytecode
and quadcode (Section \ref{sec:pretty-printing}) which is useful for
debugging analyses and deciphering their output.  The remaining
sections describe the quadcode representation along with the API of
Joeq and Chord for navigating it.  Briefly, the representation
consists of a set of classes that may be loaded
(Section \ref{sec:program}).  Each class has a set of fields and
methods, collectively called members of the class
(Section \ref{sec:class-members}).  The representation of a concrete
method (Section \ref{sec:methods}) consists of a control-flow graph
(CFG).  The representation of a CFG (Section \ref{sec:cfgs}) consists
of a set of registers and a set of basic blocks.  Each basic
block contains zero or more primitive statements called quads
(Section \ref{sec:quads}).  Finally, the most common way to traverse
all quads is discussed (Section \ref{sec:traversing}).

@@

The quadcode representation of each method is a unique object of
class \javadoc{joeq.Class.jq_Method}{joeq/Class/jq_Method.html}.  Components of the method,
most notably its control-flow graph,
can be accessed using the following API provided
by that class.

@@

The control-flow graph (CFG) of each method consists of a set of registers,
called the register factory, and a graph whose nodes are basic blocks and
whose edges denote flow of control between basic blocks.

@@

\subsection{Register Factory}

The register factory contains one register per argument of the method
(called {\it local variables}) and one register per temporary 
in the method body (called {\it stack variables}).
Temporaries include those declared by programmers as well as those
generated by Joeq.  The reason Joeq can generate temporaries is that the quadcode
representation, which is register-based, is constructed from Java
bytecode, which is stack-based; moreover, Joeq does the Static Single
Assignment (SSA) transformation by default, which introduces
temporaries to ensure that there is at most one static assignment to
any variable.  Registers corresponding to local variables are named {\tt R0}, {\tt R1}, ..., {\tt Rn},
while those corresponding to stack variables are named {\tt Tn+1}, {\tt Tn+2}, ..., {\tt Tm}.

For instance, the register factory of the main method in the example above 
has 3 registers: {\tt R0} denoting the {\tt args} argument of the method
and {\tt T1} and {\tt T2} denoting temporaries generated by Joeq.

Each register factory is a unique object
of class \javadoc{joeq.Compiler.Quad.RegisterFactory}{joeq/Compiler/Quad/RegisterFactory.html}.

\subsection{Basic Blocks}

@@

\begin{framed}
\begin{verbatim}
Operator
    |    Move, Phi, Unary, Binary, New, NewArray, MultiNewArray, 
    |--- Getstatic, Putstatic, Getfield, Putfield, ALoad, AStore,
    |    CheckCast, InstanceOf, ALength, Return
    |
    |--- Branch
    |       |--- IntIfCmp, Goto, Jsr, Ret, LookupSwitch, TableSwitch
    |
    |--- Invoke
    |       |--- InvokeVirtual, InvokeStatic, InvokeInterface
    |
    |--- Monitor
            |--- MONITORENTER, MONITOREXIT
\end{verbatim}
\end{framed}

@@

QuadVisitor qv = new QuadVisitor.EmptyVisitor() {
    public void visitMove(Quad q) { ... }
    public void visitPhi(Quad q) { ... }
    public void visitUnary(Quad q) { ... }
    ...
};
Program program = Program.g();
for (jq_Method m : program.getMethods()) {
    if (!m.isAbstract()) {
        ControlFlowGraph cfg = m.getCFG();
        ListIterator.BasicBlock it = cfg.reversePostOrderIterator();
        while (it.hasNext()) {
            BasicBlock b = it.nextBasicBlock();
            for (int i = 0; i < b.size(); i++) {
                Quad q = b.getQuad(i);
                q.accept(qv);
            }
        }
    }
}
\end{verbatim}
\end{framed}
